Sampling requirements to detect ecosystem change

### 2020-03-05 R/manuscript_draft.Rmd
Add some discussion text Mostly modified figure 2 by adding the case study time series sigma, phi, and shift size parameters.

```{r simulation_min_time_plot, fig.cap='Minimum number of samples required for 0.8 statistical power given different levels of (a) temporal variability ($\\sigma$), (b) temporal autocorrelation ($\\phi$), and (c) shift size. Random sampling was used with the default parameters are $\\sigma = 3$, $\\phi = 0.5$, and a shift size $=0.75$. The exact timing of the true changepoint varied for each simulation between time steps 30 and 70. \\label{fig:simulation_min_time}',message=F,warning=F,echo=F,eval=T}

@@

sigma <- read.csv('Output/theoretical_results_sigma.csv',header=TRUE)
plot(sigma$sigma_vec,sigma$min_time,ylab='Minimum number of samples required',xlab='sigma',pch=16,cex.lab=1.3,las=1,ylim=c(20,50),col=1,type='o')

@@

phi <- read.csv('Output/theoretical_results_phi.csv',header=TRUE)
plot(phi$phi_vec,phi$min_time,ylab=' ',xlab='phi',pch=16,cex.lab=1.3,las=1,ylim=c(20,50),col=1,type='o')

@@

shift_size <- read.csv('Output/theoretical_results_shift_size.csv',header=TRUE)
plot(shift_size$shift_size_vec,shift_size$min_time,ylab=' ',xlab='shift size',pch=16,cex.lab=1.3,las=1,ylim=c(20,50),col=1,type='o')

@@

getpath4data <- function() {
  if(Sys.getenv("USER")=="Rosalie") return("/Volumes/-/Script R/Data-off-GitHub/temporal-sampling/")
  if(Sys.getenv("USER")=="eastonwhite") return("~/Desktop/Research/soil-temporal-sampling/")
  if(Sys.getenv("USER")!="Rosalie"|Sys.getenv("USER")!="put here your USER") stop("You need to get the data and define their location.")
}
# Read data
vr <- read.delim(paste0(getpath4data(),"VAR10-10-clado.txt"))

@@

@@

@Bruel2018 processed one sample at each centimeter in a 74-centimeter sediment core. Each sample took an average of XX hours to process. We found that using an interval approach would have eliminated XX hours of sample processing or thousands of US dollars depending on labor costs.

>>>>

```{r read case study data and calculate some parameters to plot them on theoritical plot, message=FALSE, warning=FALSE, include=FALSE}
getpath4data <- function() {
  if(Sys.getenv("USER")=="Rosalie") return("/Volumes/-/Script R/Data-off-GitHub/temporal-sampling/")
  if(Sys.getenv("USER")=="eastonwhite") return("~/Desktop/Research/soil-temporal-sampling/")
  if(Sys.getenv("USER")!="Rosalie"|Sys.getenv("USER")!="put here your USER") stop("You need to get the data and define their location.")
}
# Read data
vr <- read.delim(paste0(getpath4data(),"VAR10-10-clado.txt"))

# Calculate the 3 time-series parameter needed for comparison to the theoritical results

# 1 - Calculate sigma
sd_vr <- sd(vrall$matrix$x)

# 2 - Calculate autocorrelation of real time serie
ar_vr <- acf(vrall$matrix$x, lag.max = 5, plot = FALSE)

# 3 - Calculate shift size
shift_vr <- vrall$matrix$x
# Normalize
shift_vr <- (shift_vr-min(shift_vr))/(max(shift_vr)-min(shift_vr))
# Difference between the two values before/after shift
shift_vr <- abs(shift_vr[vrall$changepoint]-shift_vr[vrall$changepoint-1])

\rosalie{I wonder whether we should show the results with random sampling for Fig. \ref{fig:simulation_min_time}: we show later that it is the worst case scenario. But maybe that's a discussion we already had and I am missing on something.

I calculated the autocorrelation at lag-1 for the case_study time series and add it to the plot here.}

```{r simulation_min_time_plot, fig.width = 7.2, fig.height= 3, fig.cap='Minimum number of samples required for 0.8 statistical power given different levels of (a) temporal variability ($\\sigma$), (b) temporal autocorrelation ($\\phi$), and (c) shift size. Random sampling was used with the default parameters are $\\sigma = 3$, $\\phi = 0.5$, and a shift size $=0.75$. The exact timing of the true changepoint varied for each simulation between time steps 30 and 70. Vertical line indicate respective parameter calculated from the case study time series. \\label{fig:simulation_min_time}',message=F,warning=F,echo=F,eval=T}

@@

sigma <- read.csv('Output/theoretical_results_sigma.csv',header=TRUE)
#plot(sigma$sigma_vec,sigma$min_time,ylab='Minimum number of samples required',xlab='sigma',pch=16,cex.lab=1.3,las=1,ylim=c(20,50),col=1,type='o')

p1 <- ggplot(sigma, aes(sigma_vec, min_time)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = ar_vr$acf[2], lty=2) +
  ylim(20,50) +
  labs(x="sigma", y= "Minimum number of samples required") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

@@

phi <- read.csv('Output/theoretical_results_phi.csv',header=TRUE)
#plot(phi$phi_vec,phi$min_time,ylab=' ',xlab='phi',pch=16,cex.lab=1.3,las=1,ylim=c(20,50),col=1,type='o')

p2 <- ggplot(phi, aes(phi_vec, min_time)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = sd_vr, lty=2) +
  ylim(20,50) +
  labs(x="phi", y= "") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

@@

shift_size <- read.csv('Output/theoretical_results_shift_size.csv',header=TRUE)
#plot(shift_size$shift_size_vec,shift_size$min_time,ylab=' ',xlab='shift size',pch=16,cex.lab=1.3,las=1,ylim=c(20,50),col=1,type='o')

p3 <- ggplot(shift_size, aes(shift_size_vec, min_time)) +
  geom_point() +
  geom_line() + 
  geom_vline(xintercept = shift_vr, lty=2) +
  ylim(20,50) +
  labs(x="phi", y= "") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

(p1 | p2 | p3)

@@

@@

Analyzing a subsample of the whole time-series inevitably leads to a lesser understanding of the phenomenon observed [@White2019d]. However, in some cases, analyzing the whole sample collected is not an option. We show here that an informed subsampling can still allow detection of critical information, such as time series changepoint.

@@

@Bruel2018 processed one sample at each centimeter in a 74-centimeter sediment core. Each sample took an average of 3 hours to process. We found that using an interval approach would have eliminated `r 3*(74-9)` hours of sample processing, or about `r round(3*(74-9)/8)` days, which is just a little over a month of work. This correspond to several thousands of US dollars depending on labor costs.

Note that the number of sample needed varies depending on the characteristics of the time series. For perspective, the autocorrelation at lag-1 in the first DCA component we used in the case study example is `r round(ar_vr$acf[2],3)`
